---
title: "comparison"
format: html
---

```{r}

log_results <- readRDS("model_outputs/logistic_predictions.rds")
rf_pred <- readRDS("model_outputs/random_forest_predictions.rds")
xgb_pred <- readRDS("model_outputs/xgboost_predictions.rds")

```


```{r}

#Create one comparison function

get_model_metrics <- function(data, model_name, class_col = ".pred_class") {

  bind_rows(
    accuracy(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col)),
    precision(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col), event_level = "first"),
    recall(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col), event_level = "first"),
    f_meas(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col), event_level = "first"),
    sens(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col), event_level = "first"),
    spec(data, truth = SeriousDlqin2yrs, estimate = !!sym(class_col), event_level = "first"),
    roc_auc(data, truth = SeriousDlqin2yrs, .pred_1, event_level = "first"),
    pr_auc(data, truth = SeriousDlqin2yrs, .pred_1, event_level = "first")
  ) |>
    mutate(model = model_name)
}

```



```{r}
#Build comparison table

log_metrics <- get_model_metrics(log_results, "Logistic Regression")
rf_metrics  <- get_model_metrics(rf_pred, "Random Forest")
xgb_metrics <- get_model_metrics(xgb_pred, "XGBoost", class_col = "final_class")


model_comparison <- bind_rows(log_metrics, rf_metrics, xgb_metrics) |>
  select(model, .metric, .estimate) |>
  pivot_wider(names_from = .metric, values_from = .estimate)|>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

model_comparison





```
#Interpretation:

Accuracy: Percentage of total predictions that were correct (high accuracy ≠ good credit model)

Precision: Of all customers predicted as defaulters, how many actually are defaulted(When the model says a borrower is risky, how often is it correct?). High precision is good.

Recall:Of all actual defaulters, how many were caught.Low recall = missed bad loans. Recall should be high.

F-measure (F1 Score) : Overall balance between catching bad borrowers and being correct when flagging them.Useful when class is imbalanced

Sensitivity: ability to detect actual defaulters.How sensitive is the model to risky customers? Same as Recall.

Specificity: Of all safe borrowers, how many were correctly identified as safe?High specificity = fewer good customers rejected.

ROC AUC:Measures model’s ability to separate: defaulters and non-defaulters.
If we randomly choose one defaulter and one safe borrower, the model ranks the defaulter as riskier 85% of the time.

PR AUC:focus on minority class performance and imbalanced datasets.
Higher is better.
How well does the model maintain precision while improving recall?

Recall + PR AUC + ROC AUC usually matter more than raw accuracy.


XGBoost provides the strongest overall performance, achieving the best discriminatory power and the highest ability to detect defaulters among all tested models.
highest recall
best ROC AUC
est PR AUC
best F1 score

```{r}

#Logistic Regression
log_cm <- conf_mat(
  log_results,
  truth = SeriousDlqin2yrs,
  estimate = .pred_class
)

log_cm


#Random Forest
rf_cm <- conf_mat(
  rf_pred,
  truth = SeriousDlqin2yrs,
  estimate = .pred_class
)

rf_cm


#XGBoost
xgb_cm <- conf_mat(
  xgb_pred,
  truth = SeriousDlqin2yrs,
  estimate = final_class
)

xgb_cm
```
```{r}

#heatmap view
autoplot(log_cm, type = "heatmap")
autoplot(rf_cm, type = "heatmap")
autoplot(xgb_cm, type = "heatmap")

```
```{r}
class(final_xgb_wf)
class(xgb_fit)

```

